### 2022-11-23 paper/paper.tex
more drafting on the pipeline steps

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
indexes \sep data pipeline \sep 
software design

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\textbf{Why index is useful, why people care about indexes}

\emph{incorporate the following in why using index: multiple pieces of
information (variables) that need to be taken into account}

Many concepts relevant to decision making cannot be directly measured,
however, they are crucial for resource allocation, early prevention, and
other operational purpose. For example, fire authorities would be
interested to quantify fire risk since bushfires can have a huge impact
on monetary loss, health, and the local ecosystem. Climatologists would
be interested in monitoring the change in global climate since
variability in atmospheric and oceanic conditions has a direct impact on
global weather and climate. Usually this concept of interest is
associated with more than one variables and these variables need to be
integrated to make decisions on the subject matter. A common approach to
quantify concepts like these is to construct an index using these
relevant variables. This allows researchers to compare the quantity of
interest across entities (i.e.~countries, regions) and also cross time.

\textbf{Define what is an index, what is not}

In this article, an index is defined as a tool to quantify a concept of
interest that does not have a direct measure. The concept of interest
doesn't have a direct measure can because it is impractical to measure
at the population level. For example, it would be nearly impossible to
include all the available stocks in the market to characterise stock
market behavior, so indexes like Dow Jones Industrial Average, S\&P 500,
and Nasdaq Composite select a representative set of stocks to measure
the overall market behavior. Also belonging to this category are the
economic indexes like the Consumer Price Index, where price changes of a
basket of items are weighted to measure inflation. The lack of direct
measure could also because the concept itself is an unobservable human
construction, rather than a physical quantity that can be measured. Many
natural hazard and social concepts falls into this category. This
includes drought indexes constructed from meteorological, agricultural,
hydrological, and social-economic variables, e.g.~Standardised
Precipitation Index (SPI) (McKee et al. 1993) and Aggregated Drought
Index (ADI) (Keyantash and Dracup 2004) among others. Social development
indexes like Human Development Index (United Nations Development
Programme 2022) and Global Liveability Index (Economist Intelligence
Unit 2019) measure various aspects of the quality of human capital and
urban life.

\emph{still need to tweak the tone a bit: ``they are called index, they
are not the index we will talk about''}

Despite many quantity having the term \emph{index} in their name, they
cannot be technically classified as indexes according to the definition
given above. The reason for these quantities to lose their index
memberships is that they are variables can be accurately measured given
the instrument precision. This includes quantities like precipitation of
the driest month or percentage of days when maximum temperature is below
10th percentile. They are measures of precipitation and percentage of
days under specific conditions (dries month, maximum temperature below
10th percentile). They are variables, or indicators, that can be used to
construct indexes but are not indexes themselves. Similarly, a set of
remote sensing indexes are not indexes, since they are measures of
electromagnetic wave reflectance. This includes Normalized Difference
Vegetation Index (NDVI) (Tucker 1979), derived from the ratio of
difference over sum on two segments in the spectrum, also called band:
near-infrared (NIR) and red. So are the ``indexes'' derived from NDVI,
e.g.~Vegetation Condition Index (Kogan 1995). Notice that this does not
exclude all the construction derived from remotes sensor variables to be
valid indexes. For example, Vegetation Drought Response Index (Brown et
al. 2008) is a valid index since it integrates climate, satellite, and
biophysical variables to quantify vegetation stress.

\textbf{What is the challenges with current index construction}

\emph{see if there is any paper describing this type of pains}

\emph{useful to reference tidy data and tidy model that makes the
workflow on modelling tidy somewhere in introduction}

Currently, index construction lacks a standardised workflow. It is often
up to researchers or research institutions to decide whether to provide
open source code on the new indexes, what would be the best user
interface for other researchers to use the new indexes, and how easily
the new indexes can be compared with other existing indexes. This makes
the computation lack transparency and indexes cumbersome to experiment
with:

\begin{itemize}
\tightlist
\item
  Researchers who wish to validate the indexes calculated from large
  institutes need to reinvent the wheels themselves since the source
  code used for computing is often not available for public consumption;
\item
  Open-source code provided by research groups has a narrow margin for
  exploring other options outside the provided;
\item
  Similar steps used by different indexes are difficult to spot since
  the design of the user interface for indexes often includes all the
  steps under a single function call; and
\item
  It is generally hard to inspect intermediate results during the index
  construction if users wish to check the output of a certain step.
\end{itemize}

\textbf{what can be done if people adopt this pipeline/ why it is
beneficial?}

This paper proposes a data pipeline for index construction. By
recognising the common steps shared by many indexes, we develop a
pipeline that breaks down index construction into multiple modules and
allow operations in various modules to be combined like building blocks
to construct indexes. The pipeline approach is general while adaptable
to most index construction. It allows indexes to be created, studied,
and compared in a structured tidy form and enables statistical analysis
of indexes to be performed easily: More specifically, it enables
researchers to 1) validate the indexes calculated from external
organisations, 2) unify various indexes under the same framework for
computing, 3) swap or adjust individual steps in the index construction
to study their contribution, 4) calculate uncertainty on indexes through
bootstrap or others, 5) enhance existing indexes through comparing and
studying their statistical properties, and finally, 6) propose new
indexes from combining different steps in existing indexes.

\textbf{who would benefit from this paper}

This work is of interest to researchers actively developing new indexes
since it encourages new indexes to be delivered in an easy-to-reproduce
design. It would also provide analysts who wish to compute a range of
indexes in their analysis a uniform interface to build relevant indexes
from raw data. For statisticians and software developing engineers, this
work frames the process of index construction in a more user-oriented
workflow and could motivate similar research for other process in
scientific computing.

The rest of the paper is structured as follows:
Section~\ref{sec-data-pipeline-in-r} reviews the concept of data
pipeline in R. The pipeline framework for index construction is
presented in
Section~\ref{sec-a-pipeline-for-building-statistical-indexes}.
Section~\ref{sec-incorporating-new-buliding-blocks-into-the-pipeline}
explains how to include a new building block in each pipeline module.
Examples are given in Section~\ref{sec-examples} to demonstrate the
index construction with the pipeline built.

\hypertarget{sec-data-pipeline-in-r}{%
\section{Data pipeline}\label{sec-data-pipeline-in-r}}

\emph{Think about if there is another word for data pipeline}

\textbf{Why you should care about pipeline}

Data pipeline is not a new concept to computing. It refers to a set of
data processing elements connected in series, where the output of one
element is the input of the next one. Wickham et al. (2009) argues that
whether made explicit or not, the pipeline has to be presented in every
graphics program. The paper also argues that breaking down graphic
rendering into steps is beneficial for understanding the implementation
and comparing between different graphic systems. The discussion on
pipeline construction is well documented in early interactive graphics
software: Buja et al. (1988), Sutherland et al. (2000), and Xie,
Hofmann, and Cheng (2014) and their pipeline steps include non-linear
transformation, variable standardization, randomization and dimension
reduction.

\textbf{What is pipeline, its underlying software design philosophy, and
how these are reflected in R}

One of the most commonly known pipeline examples is perhaps the Unix
pipeline where programs can be concatenated with \texttt{\textbar{}} to
flow the output from the last program into the next program, i.e.~

\begin{verbatim}
command 1 | command 2 | command 3 | ...
\end{verbatim}

To solve a complex problem, the Unix system builds simple programs that
do one thing well and work well together. This design is also reflected
in the tidyverse ecosystem in R. To solve a complicated data problem
using tidyverse, analysts typically build the solution using a
collection of tools from the tidyverse toolbox. The data object can flow
smoothly from one command to the next, safeguarded by the tidy data
format (Wickham 2014), which prescribes three rules on how to lay out
tabular data. The tidyverse tools also embrace a strong human-centered
design where function names are intuitive and easy to reference through
autocomplete. With the tidyverse design principle in mind, the tidymodel
suite enables analysts to build machine learning models through the data
pipeline. It includes typical tasks required in machine learning like
data resampling, feature engineering, model fitting, model tuning, and
model evaluation. An advantage of tidymodel pipeline over separate
software for individual models is that analysts no longer need to write
model-specific syntax to work with each model, but pipeline-specific
syntax that is applicable to all the models implemented in tidymodel.
This allows users to easily experiment with a collection of machine
learning models.

\textbf{Constructing indexes would also benefit from pipeline and
embracing the aforementioned design philosophy.}

In index construction, data pipeline is often presented in a workflow
diagram in the research paper to illustrate how the raw data is
transformed into the final indexes. This agrees with Wickham's argument
on the presence of the data pipeline, however, more often than not, the
pipeline is not made explicit in the software. Often the time, all the
steps are lumped into a single wrapper function, rather than being split
into smaller, modulated functions. This increases the cost of
maintaining and understanding the code base, gives analysts little
freedom to customise the indexes for specific needs, and hinders reusing
existing code for building new indexes. A pipeline approach unites a
range of indexes under a single data pipeline and analysts can compose
indexes from pipeline steps like building Legos from individual bricks.
In this workflow, analysts are not limited by indexes that have been
already proposed and can easily combine pipeline steps to compose novel
indexes. Analysis of the indexes (i.e.~calculation of uncertainty) is
also feasible by adding external code into the pipeline.

\hypertarget{sec-a-pipeline-for-building-statistical-indexes}{%
\section{A pipeline for building statistical
indexes}\label{sec-a-pipeline-for-building-statistical-indexes}}

\hypertarget{how-does-the-pipeline-constructin-of-an-index-look-like}{%
\subsection{How does the pipeline constructin of an index look
like?}\label{how-does-the-pipeline-constructin-of-an-index-look-like}}

Consider a commonly used drought index: Standardized
Precipitation-Evapotranspiration Index (SPEI) (Vicente-Serrano,
Beguería, and López-Moreno 2010). Its construction involves: 1)
calculating potential evapotranspiration (PET) from average temperature,
\texttt{tavg}, and its difference, \texttt{d}, with precipitation,
\texttt{prcp}, 2) aggregating difference series with a time scale, 3)
fitting the aggregated series with a chosen distribution to get density
values, and 4) converting the density values to normal quantiles. Under
the pipeline approach, SPEI will be constructed as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{trans\_pet}\NormalTok{(}\AttributeTok{method =}\NormalTok{ thornthwaite, }\AttributeTok{var =}\NormalTok{ tavg) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{aggregate}\NormalTok{(}\AttributeTok{scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{var =}\NormalTok{ d) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{normalise}\NormalTok{(}\AttributeTok{dist =}\NormalTok{ gamma, }\AttributeTok{fit\_method =}\NormalTok{ lmom) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{var =}\NormalTok{ .agg\_prcp)}
\end{Highlighting}
\end{Shaded}

\emph{Change the examples to two scales values, compare how indexes look
like}

Here each command corresponds to one step described above with
additional arguments specific to the step. The pipeline construction
produces the same index value as command like \texttt{spei(...)}, as
shown in Figure~\ref{fig-toy-example}, but with additional benefit:

\begin{itemize}
\tightlist
\item
  Multiple scales and multiple distributions can be fitted using the
  \texttt{c(...)} syntax to compare index values constructed from
  different parameterisations;
\item
  Intermediate results can be checked after each step; and
\item
  Additional steps and analysis can be wired into the workflow for index
  diagnostics and customised user need.
\end{itemize}

\emph{everyone is providing single command function, this is also the
same for spi}

\emph{may not worthwhile if only one index, as long as start changing}

\emph{just want to calculate index, single function is fine}

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=0.3\textheight]{../figures/toy-example-spei.png}

}

\caption{\label{fig-toy-example}Standardised Precipitation
Evapotranspiration Index (SPEI) calculated with two distribution fits in
Step 3 described above: generalised logistic (\texttt{glogist}) and
generalised extreme value (\texttt{gev}) distribution, using the
\texttt{wichita} data from the package \texttt{SPEI}.}

\end{figure}

\newpage

\hypertarget{pipeline-steps-for-constructing-indies}{%
\subsection{Pipeline steps for constructing
indies}\label{pipeline-steps-for-constructing-indies}}

An overview of the pipeline is given in Figure~\ref{fig-pipeline-steps}
to illustrate the construction from raw data to the final indexes. The
available intermediate steps include temporal processing, spatial
aggregation, variable transformation, scaling, dimension reduction,
normalising, benchmarking, and simplification. These steps are not
required to be arranged in a fixed order and not all the steps have to
appear when constructing each index. The pipeline includes modules for
operations in the spatial, temporal, and multivariate aspect of the
data, as well as modules for comparing and communicating indexes
(benchmarking and simplification). While the pipeline starts with the
raw data, steps prior to this, i.e.~define a useful concept to construct
the index, collect high quality data relevant for measuring the concept,
are also cricial to the sucess of an index. The framework proposed in
this paper assumes the objective of an index has been defined and
relevant variables have been collected and quality controlled.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{../figures/pipeline-steps.png}

}

\caption{\label{fig-pipeline-steps}Diagram of pipeline steps for index
construction.}

\end{figure}

\textbf{Raw data}

Let \(\mathbf{x}(\mathbf{s};\mathbf{t})\) denote the raw data with
spatial, temporal, and multivariate aspect: the spatial dimension
\(\mathbf{s} = (s_1, s_2, \cdots, s_n)^\prime\) is defined in the 2D
space: \(\mathbf{s} \in \mathcal{D}_s \subseteq \mathbb{R}^2\), the
temporal dimension \(\mathbf{t} = (t_1, t_2, \cdots, t_J)^\prime\) is
defined in the 1D space:
\(\mathbf{t} \in \mathcal{D}_t \subseteq \mathbb{R}\). When more than
one variable are invovled, the multivariate data can also be written as:
\(\mathbf{x}(\mathbf{s}; \mathbf{t}) = (x_1(\mathbf{s}; \mathbf{t}), x_2(\mathbf{s}; \mathbf{t}), \cdots, x_P(\mathbf{s}; \mathbf{t}))^\prime\).

\textbf{Temporal processing}

While many variables are collected separately at each time period, the
underlying phenomenon, characterised by the index, is continuous. For
example, precipitation can be measured daily or monthly, a
meteorological drought is characterised by a prelonged lack of
precipitation. Temporal proocessing is a step in the pipeline that
includes operations in the temporal dimension and it can be generally
represented by \[f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),\] where
\(\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}\) is the parameters
associated with the temporal operation and \(d_{\psi}\) is the number of
parameter of \(\psi\). The purpose of temporal aggregation can be to
aggregate variable across time or to borrow information from nearby time
periods to smooth or adjust the current value. The drought index SPI is
constructed from the aggregated precipitation with a time scale
parameter \(\psi = k\). The aggregation step in SPI can be written as
\(x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j)\)
where \(j^\prime\) is the time dimension after the aggregation.

\textbf{Spatial aggregation}

Sometimes, information from neighbouring spatial locations can also be
borrowed when constructing indexes. This involves interpolating data
from observed locations to the unobserved and aggregating data into
different levels. In the pipeline, spatial aggregation can be written as
\[g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),\] where
\(\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}\) is the spatial
operation parameters and \(d_{\theta}\) is the number of parameter of
\(\theta\). {[}GIVEN AN EXAMPLE{]}

\newpage

\textbf{Variable transformation }

Variable transformation and scaling, in the next section, are both
variable pre-processing steps to prepare the data with desire
statistical property to work with. This can be a more stablised
variance, a normal distribution that satisfies modelling assumptions, or
a data range that required by the algorithms. Variable tranformation is
a general notion of a functional transformation on the variable:
\[h_{\tau}(x(\mathbf{s};\mathbf{t})),\] where
\(\tau \in T \subseteq \mathbb{R}^{d_{\tau}}\) is the parameter in the
transformation, if any, and \(d_{\tau}\) is the number of parameter of
\(\tau\). Typical examples of variable transformation are log,
quadratic, square root, and box-cox transformation.

\textbf{Scaling}

While scaling can be seen as a specific type of variable tranformation,
it is separated into own step to make the scaling operation explicit in
the pipeline. The key difference between the two steps is that variable
transformation typically changes the shape of the data while scaling
only changes the data scale and can usually be written in the form of
\[[x(s_i;t_j) - \alpha]/\gamma.\] For example, A z-score standardisation
can be written in the above form with \(\alpha = \bar{x}(s; t)\) and
\(\gamma = \sigma(s; t)\), a min-max standardisation uses
\(\alpha = \min[x(s_i, t_j)]\) and
\(\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]\).
Figure~\ref{fig-scale-var-trans-compare} shows a collection of variable
pre-processing operations and uses color to differentiate whether the
operation is a variable transformation or a scaling step. All the
scaling operations in green results in the same shape as the original
data.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-scale-var-trans-compare-1.pdf}

}

\caption{\label{fig-scale-var-trans-compare}Comparison of scaling and
variable transformation.}

\end{figure}

\newpage

\textbf{Dimension reduction}

While multivariate information can be collected on a concept, they need
to be summarised into a single number for the final index. Dimension
reduction step
\[x_{p^*}(\mathbf{s}; \mathbf{t}) \rightarrow x_p(\mathbf{s}; \mathbf{t}),\]
where \(p^* = 1, 2, \cdots, P^*\) and \(p = 1, 2, \cdots, P\), reduces
the variable dimension from \(P\) to \(P^*\). The most commonly used
dimension reduction technique Principal Component Analysis (PCA), also
called Empirical Orthogonal Function (EOF) in earth science, can be
written as
\(x_{p^*}(\mathbf{s}; \mathbf{t}) = \sum_{p = 1}^{P}\lambda_{p}x_p(\mathbf{s};\mathbf{t})\)
where \(\lambda_p\) is the loading of the PC1, derived from maximising
the data variance given the constraint \(\sum_{p=1}^P\lambda_p^2 = 1\).

\textbf{Normalising}

The purpose of normalising is for cross-comparison using the normal
score. This step can get criticism from analysts for being unnecessarily
forcing the scale to the normal scale. This step is usually in the end
of the pipeline when the data have been processed into a univariate
series. When the data are not distributed normally, a distributed is
usually fitted before converting to the normal score via the normal
reverse CDF function: \(\Phi^{-1}(.)\):
\[\Phi^{-1}[F_{\eta}(x(\mathbf{s}; \mathbf{t}))],\] where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\).
Sometimes, to account for the monthly variation in the data, the fitting
is done separately for each month and the noramlising step can be
written as \(\Phi^{-1}[F_{\eta}^m(x(\mathbf{s};t_{j^*}))]\), where
\(j^*\) is all the indexes that satisfy \(j^* \mod 12 = m\) for each
\(m = 0, 1,\cdots, 11\).

\begin{itemize}
\tightlist
\item
  In some cases, the distibution and inverse normal can be multivariate,
  i.e.~use copula.
\end{itemize}

\textbf{Benchmarking}

Benchmarking sets a constant value to allow the constructed index to be
compared across time. Here we denote it with \(u[x(s_i, t_j)]\) where
\(u\) is a scalar of interest in the index constructed, could be a
constant or a function of the data, i.e.~mean.

\textbf{Simplification}

In public communication, indexes are often delivered in categorical
grades, along with its underlying numerical values. The simplification
step, sometimes can also be called discretisation, prescribes how the
continuous index values are converted into the discrete grades. This
process can be written with the piece-wise function:

\[
\begin{cases}
C_0 & c_1 \leq (s_i; t_j) < c_0 \\
C_1 & c_2 \leq x(s_i; t_j) < c_1 \\
C_2 & c_3 \leq x(s_i; t_j) < c_2 \\
\cdots \\
C_z & c_z \leq x(s_i; t_j)
\end{cases}
\]

where \(C_0, C_1,\cdots ,C_z\) are the categories and
\(c_0, c_1, \cdots, c_z\) are the threshold value in each category. In
SPI, drought are sorted into four categories: mild drought:
\([-0.99, 0]\); moderate drought: \([-1.49, -1]\); severe drought:
\([-1.99, -1.5]\), and extreme drought: \([-\infty, -2]\). In this case,
\(C_0, C_1, C_2, C_3\) are the drought categories: mild, moderate,
severe, and extreme drought (\(z = 3\)) and
\(c_0 =0, c_1 = -1, c_2 = -1.5, c_3 = -2\) are the cutoff value for each
class.

Discretise the continuous index into a few labelled categories. For
communicating the severity of natural hazard to general public.

uniform workflow to work with index construction.

\begin{itemize}
\tightlist
\item
  illustration
\item
  math notation
\item
  benefit of the pipeline approach

  \begin{itemize}
  \tightlist
  \item
    index diagnostic
  \item
    uncertainty
  \end{itemize}
\end{itemize}

\hypertarget{sec-incorporating-new-buliding-blocks-into-the-pipeline}{%
\section{Incorporating new buliding blocks into the
pipeline}\label{sec-incorporating-new-buliding-blocks-into-the-pipeline}}

\hypertarget{sec-examples}{%
\section{Examples}\label{sec-examples}}

\hypertarget{constructing-standardised-precipitation-index-spi}{%
\subsection{Constructing Standardised Precipitation Index
(SPI)}\label{constructing-standardised-precipitation-index-spi}}

\begin{itemize}
\tightlist
\item
  a basic workflow and congruence with results in the \texttt{SPEI} pkg
\item
  allow multiple distribution fit
\item
  allow bootstrap uncertainty
\end{itemize}

\hypertarget{calculating-spei-with-raster-data}{%
\subsection{Calculating SPEI with raster
data}\label{calculating-spei-with-raster-data}}

\hypertarget{reference}{%
\section*{Reference}\label{reference}}
\addcontentsline{toc}{section}{Reference}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-brown_vegetation_2008}{}}%
Brown, Jesslyn F., Brian D. Wardlow, Tsegaye Tadesse, Michael J. Hayes,
and Bradley C. Reed. 2008. {``The {Vegetation} {Drought} {Response}
{Index} ({VegDRI}): {A} {New} {Integrated} {Approach} for {Monitoring}
{Drought} {Stress} in {Vegetation}.''} \emph{GIScience \& Remote
Sensing} 45 (1): 16--46.
\url{https://doi.org/10.2747/1548-1603.45.1.16}.

\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.

\leavevmode\vadjust pre{\hypertarget{ref-gli}{}}%
Economist Intelligence Unit. 2019. {``The Global Liveability Index
2019.''} The Economist.
\url{https://www.cbeinternational.ca/pdf/Liveability-Free-report-2019.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-keyantash_aggregate_2004}{}}%
Keyantash, John A., and John A. Dracup. 2004. {``An Aggregate Drought
Index: {Assessing} Drought Severity Based on Fluctuations in the
Hydrologic Cycle and Surface Water Storage.''} \emph{Water Resources
Research} 40 (9). \url{https://doi.org/10.1029/2003WR002610}.

\leavevmode\vadjust pre{\hypertarget{ref-kogan_application_1995}{}}%
Kogan, F. N. 1995. {``Application of Vegetation Index and Brightness
Temperature for Drought Detection.''} \emph{Advances in Space Research},
Natural {Hazards}: {Monitoring} and {Assessment} {Using} {Remote}
{Sensing} {Technique}, 15 (11): 91--100.
\url{https://doi.org/10.1016/0273-1177(95)00079-T}.

\leavevmode\vadjust pre{\hypertarget{ref-mckee1993relationship}{}}%
McKee, Thomas B, Nolan J Doesken, John Kleist, et al. 1993. {``The
Relationship of Drought Frequency and Duration to Time Scales.''} In
\emph{Proceedings of the 8th Conference on Applied Climatology},
17:179--83. 22. Boston, MA, USA.

\leavevmode\vadjust pre{\hypertarget{ref-sutherland_orca_2000}{}}%
Sutherland, Peter, Anthony Rossini, Thomas Lumley, Nicholas Lewin-Koh,
Julie Dickerson, Zach Cox, and Dianne Cook. 2000. {``Orca: {A}
{Visualization} {Toolkit} for {High}-{Dimensional} {Data}.''}
\emph{Journal of Computational and Graphical Statistics} 9 (3): 509--29.
\url{https://www.jstor.org/stable/1390943}.

\leavevmode\vadjust pre{\hypertarget{ref-tucker_red_1979}{}}%
Tucker, Compton J. 1979. {``Red and Photographic Infrared Linear
Combinations for Monitoring Vegetation.''} \emph{Remote Sensing of
Environment} 8 (2): 127--50.
\url{https://doi.org/10.1016/0034-4257(79)90013-0}.

\leavevmode\vadjust pre{\hypertarget{ref-hdi}{}}%
United Nations Development Programme. 2022. {``Human Development Report
2021-22.''} New York. \url{http://report.hdr.undp.org}.

\leavevmode\vadjust pre{\hypertarget{ref-spei}{}}%
Vicente-Serrano, Sergio M., Santiago Beguería, and Juan I. López-Moreno.
2010. {``A {Multiscalar} {Drought} {Index} {Sensitive} to {Global}
{Warming}: {The} {Standardized} {Precipitation} {Evapotranspiration}
{Index}.''} \emph{Journal of Climate} 23 (7): 1696--1718.
\url{https://journals.ametsoc.org/view/journals/clim/23/7/2009jcli2909.1.xml}.

\leavevmode\vadjust pre{\hypertarget{ref-wickham_tidy_2014}{}}%
Wickham, Hadley. 2014. {``Tidy {Data}.''} \emph{Journal of Statistical
Software} 59 (September): 1--23.
\url{https://doi.org/10.18637/jss.v059.i10}.

\leavevmode\vadjust pre{\hypertarget{ref-wickham_plumbing_2009}{}}%
Wickham, Hadley, Michael Lawrence, Dianne Cook, Andreas Buja, Heike
Hofmann, and Deborah F. Swayne. 2009. {``The Plumbing of Interactive
Graphics.''} \emph{Computational Statistics} 24 (2): 207--15.
\url{https://doi.org/10.1007/s00180-008-0116-x}.

\leavevmode\vadjust pre{\hypertarget{ref-xie_reactive_2014}{}}%
Xie, Yihui, Heike Hofmann, and Xiaoyue Cheng. 2014. {``Reactive
{Programming} for {Interactive} {Graphics}.''} \emph{Statistical
Science} 29 (2): 201--13.
\url{https://www.jstor.org/stable/43288470?seq=1}.

\end{CSLReferences}

\end{document}